
---
title: "HRV_Data_Analysis"
author: "Taliah Soleymani"
date: "2024-07-05"
output: html_document
---

## Load relevant packages

```{r}
# Assign libs with the packages needed by using c()
libs <- c("tidyr", "dplyr", "ggplot2", "ggrepel", "tibble", "rmarkdown", "pheatmap", "colorspace", "colormap", "reshape2", "data.table", "scales", "tidyverse", "stringr", "readxl", "patchwork", "lmertest")

# lapply() function is useful for performing operations on list objects
lapply(libs, require, character.only = TRUE)

# rm() removes object from the environment
rm(libs)
```

## Set directory

Make sure to save the files to a central location and set the directory to that location.

```{r}
setwd("/Users/soleymanitl/Documents/HRV")
```

### Set the directory for saving plots and output
```{r}
# Define the base directory as a variable
base_dir <- "/Users/soleymanitl/Documents/HRV/"

# Use the base directory to construct plot_dir and output_dir
plot_dir <- file.path(base_dir, "Test_Subjects_Analysis_Output")
output_dir <- file.path(base_dir, "Test_Subjects_Analysis_Output")
```

## Load HRV_Data_Loop_Test and MADRS_Sample

```{r}
# Construct the paths using the base directory
#define the path of the sample files or your own data
hrv_data_path <- file.path(base_dir, "Test_Subjects_Physio_Output", "HRV_Data_Loop_Test.csv")
madrs_data_path <- file.path(base_dir, "Data", "MADRS_Sample.csv")

# Load the data using the constructed paths
HRVData <- read.csv(hrv_data_path)
head(HRVData)

MADRS <- read.csv(madrs_data_path)
head(MADRS)

#Load here HRV_ConfoundingFactors_Sample or your own file of confounding factors 
CF_data_path <- file.path(base_dir, "Data", "HRV_ConfoundingFactors_Sample.csv")

# Load the data using the constructed paths
CF <- read.csv(CF_data_path)
head(CF)

```

- The `HRV_Data_Loop_Test.csv` spreadsheet contains the heart rate (HR) and heart rate variability (HRV) measures.
- The `MADRS_Sample.csv` spreadsheet contains the MADRS scores (mood ratings) pre and post-infusion.
- Note: the data has been changed and should not be used to test out hypotheses. Additionally, many of these plots will not look complete due to the sample being only of three patients. 

## 1. Organize HRV_Data_Loop_Test and MADRS

### 1.a. Subset and Modify Dataframes

#### Subset HRV_Data_Loop_Test to HRVData

```{r}
HRVData <- subset(HRVData, select = c("sub", "ses", "run", "PPG_Rate_Mean", "HRV_SDANN1", "HRV_SDNNI1", "HRV_RMSSD", "HRV_pNN50", "HRV_LFHF", "HRV_LFn", "HRV_HFn"))
head(HRVData)
```

#### Modify Column Names

```{r}
HRVData <- HRVData %>%
  rename(Subject_number = sub, Session_date = ses)
head(HRVData)

MADRS <- MADRS %>%
  rename(Subject_number = Subject_ID)
head(MADRS)
```

#### Format Columns

```{r}
# HRVData formatting
HRVData$Subject_number <- substring(HRVData$Subject_number, nchar(as.character(HRVData$Subject_number)) - 2)
HRVData$Session_date <- substring(HRVData$Session_date, nchar(as.character(HRVData$Session_date)) - 7)
head(HRVData)

# MADRS formatting
MADRS$Subject_number <- substring(MADRS$Subject_number, nchar(as.character(MADRS$Subject_number)) - 2)
MADRS$Session_date <- substring(MADRS$Session_date, nchar(as.character(MADRS$Session_date)) - 7)
head(MADRS)
```

### 1.b. Label Scans

#### Label Scans

```{r}
table(HRVData$run)

HRVData$Timepoint <- ifelse(grepl("run-001", HRVData$run), "Infusion_Baseline",
                             ifelse(grepl("run-01", HRVData$run), "Infusion_Baseline",
                                    ifelse(grepl("run-101", HRVData$run), "Infusion_Beginning",
                                           ifelse(grepl("run-201", HRVData$run), "Infusion_End", NA))))
head(HRVData)
```
### 1.c. Prepare MADRS and HRVData to Merge 
#### Add Timepoint to the MADRS

```{r}
MADRS <- within(MADRS, Timepoint <- ifelse(grepl("pre", MADRS$Interval), "Infusion_Baseline",
                                                         ifelse(grepl("post", MADRS$Interval), "Infusion_End", NA)))
```


#### Add Drug Column to HRVData

```{r}
HRVData <- merge(HRVData, MADRS[, c("Subject_number", "Session_date", "Drug", "Infusion")], 
                  by = c("Subject_number", "Session_date"))
head(HRVData)
```

#### Merge HRVData and MADRS into HRV_MADRS

```{r}
HRV_MADRS <- merge(MADRS, HRVData[, c("Subject_number", "Session_date", "Timepoint", "PPG_Rate_Mean", "HRV_SDANN1", "HRV_SDNNI1", "HRV_RMSSD", "HRV_pNN50", "HRV_LFHF", "HRV_LFn", "HRV_HFn")], 
                   by = c("Subject_number", "Session_date", "Timepoint"))

# Display the result
print(HRV_MADRS)
```

## 2. EDA - Visualizing HRV, Mood Ratings, and KET vs. PBO

### 2.a. Distribution of HR, BMI, Age

#### Histogram of HR

```{r}
HRVData %>% ggplot(aes(x = PPG_Rate_Mean)) +
  geom_histogram(binwidth = 0.5)
```

#### Filter Out Abnormal HRs

```{r}
HRVData %>% filter(PPG_Rate_Mean > 40) %>% filter(PPG_Rate_Mean < 110) %>% ggplot(aes(x = PPG_Rate_Mean)) +
  geom_histogram(binwidth = 0.5)
```

#### Merge HRV_MADRS with CF HRV_ConfoundingFactors_Sample

```{r}

HRV_MADRS$Age <- CF$Age[match(HRV_MADRS$Subject_number, CF$Subject_number)]
HRV_MADRS$Sex <- CF$Sex[match(HRV_MADRS$Subject_number, CF$Subject_number)]
HRV_MADRS$BMI <- CF$BMI[match(HRV_MADRS$Subject_number, CF$Subject_number)]

head(HRV_MADRS)
```

#### Histogram of BMI

```{r}
CF %>% ggplot(aes(x = BMI)) +
  geom_histogram(binwidth = 0.5)
```

#### Histogram of Age

```{r}
CF %>% ggplot(aes(x = Age)) +
  geom_histogram(binwidth = 0.5)
```

### 2.b. Correlation Between Confounding Factors and HR

#### Scatterplot: Age vs. HR

```{r}
HRV_MADRS %>% filter(Timepoint %in% c("Infusion_Baseline")) %>% ggplot(aes(Age, PPG_Rate_Mean)) +
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE, color = "blue")
```

#### Scatterplot: BMI vs. HR

```{r}
HRV_MADRS %>% filter(Timepoint %in% c("Infusion_Baseline")) %>% ggplot(aes(BMI, PPG_Rate_Mean)) +
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE, color = "blue")
```

#### Scatterplot: Sex vs. HR

```{r}
table(HRV_MADRS$Sex)

Sex_Average_HR_Baseline <- HRV_MADRS %>%
  filter(Timepoint %in% c("Infusion_Baseline")) %>%
  group_by(Sex) %>%
  summarise(avg_PPG_Rate_Mean = mean(PPG_Rate_Mean))

HRV_MADRS %>% filter(Timepoint %in% c("Infusion_Baseline")) %>% ggplot(aes(Sex, PPG_Rate_Mean)) +
  geom_point() + 
  geom_point(data = Sex_Average_HR_Baseline, aes(x = Sex, y = avg_PPG_Rate_Mean), color = "red", size = 3) + 
  labs(title = "Scatterplot with Average For Baseline", x = "Gender", y = "PPG Rate Mean")
```

### 2.c. Drug and Mood Effect

#### Mood Ratings Post-Infusion

```{r}
HRV_MADRS %>%
  filter(Timepoint %in% c("Infusion_End")) %>%
  ggplot(aes(Drug, MADRS_Total, group = Subject_number, color = as.factor(Subject_number))) +

  
  geom_point() + 
  geom_line()
```

#### Mood Ratings Pre- and Post-Infusion

```{r}
HRV_MADRS %>%
  ggplot(aes(x = Drug, y = MADRS_Total)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.2) +
  facet_wrap(~Timepoint)
```

### 2.d. Effect of Randomization (AB vs BA)

#### Assign Randomization Pattern to Scans

```{r}
HRV_MADRS_recoded <- HRV_MADRS %>%
  mutate(Randomization = case_when(
    Infusion == "1" & Drug == "A" ~ "AB",
    Infusion == "2" & Drug == "A" ~ "BA",
    Infusion == "2" & Drug == "B" ~ "AB",
    Infusion == "1" & Drug == "B" ~ "BA",
    TRUE ~ NA_character_))
head(HRV_MADRS_recoded)
table(HRV_MADRS_recoded$Randomization)
table(HRV_MADRS_recoded$Infusion)
```


#### Facet Plot of MADRS_Totals Pre- and Post-Infusion

```{r}
HRV_MADRS_recoded %>%
  filter(!(Subject_number %in% c("108", "110", "114"))) %>%
  ggplot(aes(x = Infusion, y = MADRS_Total)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.2) +
  facet_grid(Timepoint ~ Randomization)
```

### 2.e. Difference in Measures During vs. Baseline

#### Calculate the Difference for All Measures

```{r}
HRV_MADRS_Result <- HRV_MADRS_recoded %>%
  filter(Timepoint %in% c("Infusion_End", "Infusion_Baseline")) %>% 
  group_by(Subject_number, Session_date, Infusion, Drug, Randomization, Sex, BMI, Age) %>%
  arrange(Subject_number, Session_date, Infusion, Timepoint) %>%
  reframe(
    MADRS_Total_diff = MADRS_Total - lag(MADRS_Total),
    PPG_Rate_Mean_diff = PPG_Rate_Mean - lag(PPG_Rate_Mean),
    HRV_SDANN1_diff = HRV_SDANN1 - lag(HRV_SDANN1),
    HRV_SDNNI1_diff = HRV_SDNNI1 - lag(HRV_SDNNI1),
    HRV_RMSSD_diff = HRV_RMSSD - lag(HRV_RMSSD),
    HRV_pNN50_diff = HRV_pNN50 - lag(HRV_pNN50),
    HRV_LFHF_diff = HRV_LFHF - lag(HRV_LFHF),
    HRV_LFn_diff = HRV_LFn - lag(HRV_LFn),
    HRV_HFn_diff = HRV_HFn - lag(HRV_HFn)) %>% na.omit()

# Print the result
head(HRV_MADRS_Result)
```

#### Difference in Mood Rating

```{r}
HRV_MADRS_Result %>%
  ggplot(aes(x = Drug, y = MADRS_Total_diff)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.2)
```

#### Difference in Mood Ratings with Randomization Effect

```{r}
HRV_MADRS_Result %>%
  ggplot(aes(x = Infusion, y = MADRS_Total_diff)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.2) +
  facet_wrap(~Randomization)
```
#### Merge difference in MADRS score and HRV data 

```{r}
MADRS_diff_HRV <- merge(HRVData, HRV_MADRS_Result[c("Subject_number", "Session_date", "MADRS_Total_diff")], 
                     by = c("Subject_number", "Session_date"), all.x = TRUE)
```


### 2.f. Drug Effect on HRV Measures

#### HRV Measures Post-Infusion

```{r}
# Different measures PPG_Rate_Mean, HRV_SDANN1, HRV_RMSSD, HRV_pNN50, HRV_LFHF, HRV_LFn, HRV_HFn
PPG_Rate_Mean_Plot1 <- HRV_MADRS %>%
  filter(Timepoint %in% c("Infusion_End")) %>%
  ggplot(aes(Drug, PPG_Rate_Mean, group = Subject_number, color = as.factor(Subject_number))) +  
  geom_point() + 
  geom_line()

HRV_SDANN1_Plot1 <- HRV_MADRS %>%
  filter(Timepoint %in% c("Infusion_End")) %>%
  ggplot(aes(Drug, HRV_SDANN1, group = Subject_number, color = as.factor(Subject_number))) +  
  geom_point() + 
  geom_line()

HRV_RMSSD_Plot1 <- HRV_MADRS %>%
  filter(Timepoint %in% c("Infusion_End")) %>%
  ggplot(aes(Drug, HRV_RMSSD, group = Subject_number, color = as.factor(Subject_number))) +  
  geom_point() + 
  geom_line()

HRV_pNN50_Plot1 <- HRV_MADRS %>%
  filter(Timepoint %in% c("Infusion_End")) %>%
  ggplot(aes(Drug, HRV_pNN50, group = Subject_number, color = as.factor(Subject_number))) +  
  geom_point() + 
  geom_line()

HRV_HFn_Plot1 <- HRV_MADRS %>%
  filter(Timepoint %in% c("Infusion_End")) %>%
  ggplot(aes(Drug, HRV_HFn, group = Subject_number, color = as.factor(Subject_number))) +  
  geom_point() + 
  geom_line()

HRV_LFn_Plot1 <- HRV_MADRS %>%
  filter(Timepoint %in% c("Infusion_End")) %>%
  ggplot(aes(Drug, HRV_LFn, group = Subject_number, color = as.factor(Subject_number))) +  
  geom_point() + 
  geom_line()

HRV_LFHF_Plot1 <- HRV_MADRS %>%
  filter(Timepoint %in% c("Infusion_End")) %>%
  ggplot(aes(Drug, HRV_LFHF, group = Subject_number, color = as.factor(Subject_number))) +  
  geom_point() + 
  geom_line()

Combined_HRV_Plot1 <- HRV_SDANN1_Plot1 + HRV_RMSSD_Plot1 + HRV_pNN50_Plot1 + HRV_HFn_Plot1 + HRV_LFn_Plot1 + HRV_LFHF_Plot1

# Visualize all the plots side by side
Combined_HRV_Plot1

# Add the Mood Plot 
Mood_Plot1 <- HRV_MADRS %>%
  filter(Timepoint %in% c("Infusion_End")) %>%
  ggplot(aes(Drug, MADRS_Total, group = Subject_number, color = as.factor(Subject_number))) +  
  geom_point() + 
  geom_line()

Combined_HRV_Mood_Plot1 <- Mood_Plot1 + PPG_Rate_Mean_Plot1 + HRV_RMSSD_Plot1 + HRV_LFn_Plot1 + HRV_LFHF_Plot1
Combined_HRV_Mood_Plot1
```

#### Boxplots of HRV Measures

```{r}
# Different measures PPG_Rate_Mean, HRV_SDANN1, HRV_RMSSD, HRV_pNN50, HRV_LFHF, HRV_LFn, HRV_HFn

# PPG_Rate_Mean
HRV_MADRS_recoded %>%
  filter(Timepoint %in% c("Infusion_End")) %>%
  ggplot(aes(x = Drug, y = PPG_Rate_Mean)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.2)

# HRV_RMSSD
HRV_MADRS_recoded %>%
  filter(Timepoint %in% c("Infusion_End")) %>%
  ggplot(aes(x = Drug, y = HRV_RMSSD)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.2)

# HRV_LFn
HRV_MADRS_recoded %>%
  filter(Timepoint %in% c("Infusion_End")) %>%
  ggplot(aes(x = Drug, y = HRV_LFn)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width =

 0.2)

# HRV_HFn
HRV_MADRS_recoded %>%
  filter(Timepoint %in% c("Infusion_End")) %>%
  ggplot(aes(x = Drug, y = HRV_HFn)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.2)
```


## 3. The Final Plots

### 3.a. Comparing Baseline and Post-infusion Ratings

```{r}
library(ggplot2)
library(dplyr)

HRV_MADRS <- HRV_MADRS %>%
  mutate(Drug_Label = ifelse(Drug == "A", "KET", "PBO"),
         Adjusted_Timepoint = ifelse(Timepoint == "Infusion_Baseline", "Baseline", "After Infusion"),
         Adjusted_Timepoint = factor(Adjusted_Timepoint, levels = c("Baseline", "After Infusion")))

plot <- HRV_MADRS %>%
  ggplot(aes(x = Adjusted_Timepoint, y = MADRS_Total, color = Drug_Label)) +
  geom_boxplot(outlier.shape = NA, size = 0.5) +  # Increased line thickness
  geom_jitter(width = 0.2, size = 2) +  # Increased dot size
  facet_wrap(~Drug_Label) +
  labs(x = "Time Point", y = "Mood Rating") +  # Labels for X and Y axes
  scale_color_manual(values = c("KET" = "#4E81BD", "PBO" = "#595959"), name = "Drug") +  # Updated color for PBO
  theme_light() +  # Applying classic theme
  theme(panel.grid = element_blank(), text = element_text(size = 28))  # Removing grid lines and setting font size

print(plot)
ggsave(filename = file.path(plot_dir, "HRV_KET_PBO_Mood_before_after.png"), plot = plot, width = 10, height = 6, dpi = 300)
```

### 3.b. HR Over Time


```{r}

# Filter out rows with NAs in the Timepoint column
HRVData2 <- HRVData[complete.cases(HRVData$Timepoint), ]

# Define the factor levels for Timepoint
timepoint_levels <- c("Infusion_Baseline", "Infusion_Beginning", "Infusion_End")

# Replace "A" with "KET" and "B" with "PBO" in the Drug column
HRVData2$Drug <- ifelse(HRVData2$Drug == "A", "KET", "PBO")

# Calculate the average for each drug and time point
avg_data <- HRVData2 %>%
  group_by(Drug, Timepoint) %>%
  summarize(avg_PPG_Rate_Mean = mean(PPG_Rate_Mean),
            se = sd(PPG_Rate_Mean) / sqrt(n()),
            .groups = 'drop')  # Add .groups argument to avoid the grouping issue

# Plot for both drugs overlapped
combined_plot <- ggplot(avg_data %>% filter(!is.na(avg_PPG_Rate_Mean)), 
                        aes(x = factor(Timepoint, levels = timepoint_levels), y = avg_PPG_Rate_Mean, color = Drug, group = Drug)) +
  geom_line(size = 1, show.legend = TRUE) +
  geom_errorbar(aes(ymin = avg_PPG_Rate_Mean - se, ymax = avg_PPG_Rate_Mean + se), width = 0.25, show.legend = TRUE) +
  labs(x = "Time Point", y = "HR (BPM)", color = "Drug") +
  scale_x_discrete(labels = c("Infusion_Baseline" = "Baseline", "Infusion_Beginning" = "Beginning of Infusion", "Infusion_End" = "End of Infusion")) +
  scale_color_manual(values = c("KET" = "#4E81BD", "PBO" = "#595959"), name = "Drug", labels = c("KET" = "KET", "PBO" = "PBO")) +
  theme_light() +
  theme(panel.grid = element_blank(), text = element_text(size = 28)) +
  coord_cartesian(clip = "off")  # Prevent clipping of long x-axis labels

print(combined_plot)
ggsave(filename = "HRV_averageHR_KET_PBO_combined.png", plot = combined_plot, width = 12, height = 6, dpi = 300)

```


### 3.c. HRV Over Time


```{r}
# Calculate the average for each drug and time point
avg_data <- HRVData2 %>%
  group_by(Drug, Timepoint) %>%
  summarize(avg_HRV_RMSSD = mean(HRV_RMSSD),
            se = sd(HRV_RMSSD) / sqrt(n()),
            .groups = 'drop')  # Add .groups argument to avoid the grouping issue

# Plot for both drugs overlapped
combined_plot <- ggplot(avg_data %>% filter(!is.na(avg_HRV_RMSSD)), 
                        aes(x = factor(Timepoint, levels = timepoint_levels), y = avg_HRV_RMSSD, color = Drug, group = Drug)) +
  geom_line(size = 2, show.legend = TRUE) +  # Increased line size to 2
  geom_errorbar(aes(ymin = avg_HRV_RMSSD - se, ymax = avg_HRV_RMSSD + se), width = 0.25, show.legend = TRUE) +
  labs(x = "Time Point", y = "HRV (RMSSD)", color = "Drug") +
  scale_x_discrete(labels = c("Infusion_Baseline" = "Baseline", "Infusion_Beginning" = "Beginning of Infusion", "Infusion_End" = "End of Infusion")) +
  scale_color_manual(values = c("KET" = "#4E81BD", "PBO" = "#595959"), name = "Drug", labels = c("KET" = "KET", "PBO" = "PBO")) +
  theme_light() +
  theme(panel.grid = element_blank(), text = element_text(size = 28)) +
  coord_cartesian(clip = "off")  # Prevent clipping of long x-axis labels

print(combined_plot)
ggsave(filename = "HRV_averageHRV_KET_PBO_combined.png", plot = combined_plot, width = 12, height = 6, dpi = 300)
```

### 3.d. Effect of Placebo on HRV

```{r}
# Create a new grouping variable for infusion groups
HRVData3 <- HRVData2 %>%
  mutate(Infusion_Group = case_when(
    Infusion == 1 ~ "1st Infusion",
    Infusion == 2 ~ "2nd Infusion",
    TRUE ~ NA_character_
  )) %>%
  filter(Timepoint %in% c("Infusion_Baseline", "Infusion_Beginning")) %>%
  filter(Drug == "PBO") %>% # Make sure Drug is either PBO or whatever your placebo is labeled as
  drop_na()  # Remove rows with NA values in any column

# Ensure there is data for both infusion groups
table(HRVData3$Infusion_Group)

# Plot the data
plot <- ggplot(HRVData3, aes(Timepoint, HRV_RMSSD, group = Subject_number)) +
  geom_jitter(width = 0, height = 0.2, size = 4, color = "black", show.legend = FALSE) +  # Add jittered black points with only vertical jitter
  geom_line(color = "gray") +  # Add lines connecting dots for each subject
  labs(x = "Time point", y = "HRV (RMSSD)") +
  scale_x_discrete(labels = c("Infusion_Baseline" = "Baseline",
                              "Infusion_Beginning" = "Beginning\nof Infusion"),
                   expand = c(0.4, 0.4)) +  # Add more space between time points
  theme_light() +  # Apply light theme
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        text = element_text(size = 28)) +  # Set font size to 28
  coord_cartesian(clip = "on") +  # Label axes
  facet_grid(~Infusion_Group)  # Facet by Infusion Group (1st and 2nd Infusion)

# Display the plot
print(plot)

ggsave(filename = "Effect_Placebo_HRV.png", plot = combined_plot, width = 12, height = 6, dpi = 300)
```

### 3.e. Correlation Between Change in Mood Rating and HR Level at Baseline

```{r}
library(ggplot2)
library(dplyr)

# Assuming MADRS_diff_HRV is your merged dataset containing columns MADRS_Total_diff, PPG_Rate_Mean, and Timepoint

# Filter the data for Timepoint "Infusion Baseline"
filtered_data <- filter(MADRS_diff_HRV, Timepoint == "Infusion_Baseline")

# Create the scatter plot with a trendline
plot_d <- ggplot(filtered_data, aes(x = MADRS_Total_diff, y = PPG_Rate_Mean)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "#4E81BD") +  # Add trendline
  labs(x = "Change in Mood Rating", y = "HR (BPM)", title = "Mood Rating Difference VS HR at Baseline") +
  theme_light() +  # Applying classic theme
  theme(panel.grid = element_blank(), text = element_text(size = 28)) +  # Setting font size to 28
  coord_cartesian(clip = "off")  # Prevent clipping of long x-axis labels
plot_d
ggsave("HR_mood_diff_baseline.png", plot_d, width = 12, height = 6, dpi = 300)  # Increased width

```

### 3.g. Correlation Between Change in Mood Rating and HR Level at the End of Infusion
```{r}
library(ggplot2)
library(dplyr)

# Assuming MADRS_diff_HRV is your merged dataset containing columns MADRS_Total_diff, PPG_Rate_Mean, and Timepoint

# Filter the data for Timepoint "Infusion_End"
filtered_data <- filter(MADRS_diff_HRV, Timepoint == "Infusion_End")

# Create the scatter plot with a trendline
plot_d <- ggplot(filtered_data, aes(x = MADRS_Total_diff, y = PPG_Rate_Mean)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "#4E81BD") +  # Add trendline
  labs(x = "Change in Mood Rating", y = "HR (BPM)", title = "Mood Rating Difference VS HR During Infusion") +
  theme_light() +  # Applying classic theme
  theme(panel.grid = element_blank(), text = element_text(size = 28)) +  # Setting font size to 28
  coord_cartesian(clip = "off")  # Prevent clipping of long x-axis labels

ggsave("HR_mood_diff_end.png", plot_d, width = 12, height = 6, dpi = 300)  # Increased width

```


## 4. The Model

### 4.a. Mood_rating Model

#### Prepare Data for Model

```{r}
# Make data frame that aligns with model 
Model_mood <- subset(HRV_MADRS_recoded, Timepoint == "Infusion_During", select = c("Subject_number", "Drug", "Infusion", "Session_date", "Mood_rating"))
head(Model_mood)

# Add Average Baseline 
average_mood_baseline <- HRV_MADRS_recoded %>%
  filter(Timepoint == "Infusion_Baseline") %>%
  group_by(Subject_number) %>%
  summarize(average_mood_baseline = mean(Mood_rating, na.rm = TRUE))

Model_mood <- left_join(Model_mood, average_mood_baseline, by = "Subject_number")
head(Model_mood)

# Add column Mood_Baseline
HRV_MADRS_renamed <- HRV_MADRS_recoded %>% 
  filter(Timepoint == "Infusion_Baseline") %>% 
  select(Subject_number, Session_date, Mood_rating) %>%
  rename(Mood_baseline = Mood_rating)

Model_mood <- left_join(Model_mood, HRV_MADRS_renamed, by = c("Subject_number", "Session_date"))
head(Model_mood)
```

#### Fit Mood_rating Model

```{r}
library(Matrix)
library(lme4)

# Relevel the "Drug" variable to set "DrugB" as the reference level
Model_mood$Drug <- factor(Model_mood$Drug)
Model_mood$Drug <- relevel(Model_mood$Drug, ref = "B")

# Fit the mixed-effects model
M1 <- lmer(Mood_rating ~ Drug + Mood_baseline + average_mood_baseline + Infusion_Recoded + (1 | Subject_number), data = Model_mood)

# Save the output as PDF
output_file <- file.path(output_dir, "Mood_rating_model_output.pdf")
pdf(output_file)
print(summary(M1))
emm <- emmeans(M1, "Drug")
pairwise <- pairs(emm, adjust = "tukey")
print(pairwise)
dev.off()
```

### 4.b. HR Model

#### Prepare Data for Model

```{r}
# Make data frame that aligns with the model
Model_HR <- subset(HRV_MADRS_recoded, Timepoint == "Infusion_During", select = c("Subject_number", "Drug", "Infusion_Recoded", "Session_date", "PPG_Rate_Mean"))
head(Model_HR)

# Add Average Baseline 
average_PPG_Rate_Mean <- HRV_MADRS_recoded %>%
  filter(Timepoint == "Infusion_Baseline") %>%
  group_by(Subject_number) %>%
  summarize(average_PPG_Rate_Mean = mean(PPG_Rate_Mean, na.rm = TRUE))

Model_HR <- left_join(Model_HR, average_PPG_Rate_Mean, by = "Subject_number")
head(Model_HR)

# Add column PPG_Rate_Baseline
HRV_MADRS_renamed <- HRV_MADRS_recoded %>% 
  filter(Timepoint == "Infusion_Baseline") %>% 
  select(Subject_number, Session_date, PPG_Rate_Mean) %>%
  rename(PPG_Rate_Baseline = PPG_Rate_Mean)

Model_HR <- left_join(Model_HR, HRV_MADRS_renamed, by = c("Subject_number", "Session_date"))
head(Model_HR)
```

#### Fit HR Model

```{r}
library(Matrix)
library(lme4)

# Relevel the "Drug" variable to set "DrugB" as the reference level
Model_HR$Drug <- factor(Model_HR$Drug)
Model_HR$Drug <- relevel(Model_HR$Drug, ref = "B")

# Fit the mixed-effects model
M2 <- lmer(PPG_Rate_Mean ~ Drug + PPG_Rate_Baseline + average_PPG_Rate_Mean + Infusion + (1 | Subject_number), data = Model_HR)

# Save the output as PDF
output_file <- file.path(output_dir, "HR_model_output.pdf")
pdf(output_file)
print(summary(M2))
emm <- emmeans(M2, "Drug")
pairwise <- pairs(emm, adjust = "tukey")
print(pairwise)
dev.off()
```

### 4.c. HRV_RMSSD Model

#### Prepare Data for Model

```{r}
# Make data frame that aligns with the model
Model_RMSSD <- subset(HRV_MADRS_recoded, Timepoint == "Infusion_During", select = c("Subject_number", "Drug", "Infusion", "Session_date", "HRV_RMSSD"))
head(Model_RMSSD)

# Add Average Baseline 
average_HRV_RMSSD <- HRV_MADRS_recoded %>%
  filter(Timepoint == "Infusion_Baseline") %>%
  group_by(Subject_number) %>%
  summarize(average_HRV_RMSSD = mean(HRV_RMSSD, na.rm = TRUE))

Model_RMSSD <- left_join(Model_RMSSD, average_HRV_RMSSD, by = "Subject_number")
head(Model_RMSSD)

# Add column HRV_RMSSD_Baseline
HRV_MADRS_renamed <- HRV_MADRS_recoded %>% 
  filter(Timepoint == "Infusion_Baseline") %>% 
  select(Subject_number, Session_date, HRV_RMSSD) %>%
  rename(HRV_RMSSD_Baseline = HRV_RMSSD)

Model_RMSSD <- left_join(Model_RMSSD, HRV_MADRS_renamed, by = c("Subject_number", "Session_date"))
head(Model_RMSSD)
```

#### Fit HRV_RMSSD Model

```{r}
library(Matrix)
library(lme4)

# Relevel the "Drug" variable to set "DrugB" as the reference level
Model_RMSSD$Drug <- factor(Model_RMSSD$Drug)
Model_RMSSD$Drug <- relevel(Model_RMSSD$Drug, ref = "B")

# Fit the mixed-effects model
M3 <- lmer(HRV_RMSSD ~ Drug + HRV_RMSSD_Baseline + average_HRV_RMSSD + Infusion + (1 | Subject_number), data = Model_RMSSD)

# Save the output as PDF
output_file <- file.path(output_dir, "HRV_RMSSD_model_output.pdf")
pdf(output_file)
print(summary(M3))
emm <- emmeans(M3, "Drug*Timepoint")
pairwise <- pairs(emm, adjust = "tukey")
print(pairwise)
dev.off()
```

### 4.d. HRV_LFn Model

#### Prepare Data for Model

```{r}
# Make data frame that aligns with the model
Model_LFn <- subset(HRV_MADRS_recoded, Timepoint == "Infusion_During", select = c("Subject_number", "Drug", "Infusion", "Session_date", "HRV_LFn"))
head(Model_LFn)

# Add Average Baseline 
average_HRV_LFn <- HRV_MADRS_recoded %>%
  filter(Timepoint == "Infusion_Baseline") %>%
  group_by(Subject_number) %>%
  summarize(average_HRV_LFn = mean(HRV_LFn, na.rm = TRUE))

Model_LFn <- left_join(Model_LFn, average_HRV_LFn, by = "Subject_number")
head(Model_LFn)

# Add column HRV_LFn_Baseline
HRV_MADRS_renamed <- HRV_MADRS_recoded %>% 
  filter(Timepoint == "Infusion_Baseline") %>% 
  select(Subject_number, Session_date, HRV_LFn) %>%
  rename(HRV_LFn_Baseline = HRV_LFn)

Model_LFn <- left_join(Model_LFn, HRV_MADRS_renamed, by = c("Subject_number", "Session_date"))
head(Model_LFn)
```

#### Fit HRV_LFn Model

```{r}
library(Matrix)
library(lme4)

# Relevel the "Drug" variable to set "DrugB" as the reference level
Model_LFn$Drug <- factor(Model_LFn$Drug)
Model_LFn$Drug <- relevel(Model_LFn$Drug, ref = "B")

# Fit the mixed-effects model
M4 <- lmer(HRV_LFn ~ Drug + HRV_LFn_Baseline + average_HRV_LFn + Infusion + (1 | Subject_number), data = Model_LFn)

# Save the output as PDF
output_file <- file.path(output_dir, "HRV_LFn_model_output.pdf")
pdf(output_file)
print(summary(M4))
emm <- emmeans(M4, "Drug")
pairwise <- pairs(emm, adjust = "tukey")
print(pairwise)
dev.off()

```

### 4.e. Mood ~ Drug*PPG_Rate_Mean Model

#### Prepare Data for Model

```{r}
# Join Model_mood with relevant columns from Model_HR
common_cols <- c("Subject_number", "Drug", "Infusion_Recoded", "Session_date")

Combined_Model <- left_join(Model_mood, 
                            Model_HR %>% select(one_of(common_cols, "PPG_Rate_Mean", "average_PPG_Rate_Mean", "PPG_Rate_Baseline")),
                            by = common_cols)

# Check the combined data frame
head(Combined_Model)
```

#### Fit Mood ~ Drug*PPG_Rate_Mean Model

```{r}
library(Matrix)
library(lme4)

# Relevel the "Drug" variable to set "DrugB" as the reference level
Combined_Model$Drug <- factor(Combined_Model$Drug)
Combined_Model$Drug <- relevel(Combined_Model$Drug, ref = "B")

# Fit the mixed-effects model
M5 <- lmer(Mood_rating ~ Drug * PPG_Rate_Mean + Mood_baseline + average_mood_baseline + Infusion + (1 | Subject_number), data = Combined_Model)

# Save the output as PDF
output_file <- file.path(output_dir, "Mood_Drug_PPG_Rate_Mean_model_output.pdf")
pdf(output_file)
print(summary(M5))
emm <- emmeans(M5, "Drug")
pairwise <- pairs(emm, adjust = "tukey")
print(pairwise)
dev.off()

# Plot effects
library(effects)
pdf(file.path(output_dir, "Mood_Drug_PPG_Rate_Mean_model_effects.pdf"))
plot(allEffects(M5))
plot(effect(term = "Drug", mod = M5))
dev.off()
```
```
